Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-06-09 doc/thesis/methods_results.Rnw
Updated grid search results

{\color{red}\sout{The accuracy of grid search estimates largely paralleled that
of the \gls{kSVR} classifiers.}}\Cref{fig:gridest} shows point estimates and
95\% highest density intervals for each of the \gls{BA} parameters, for one
replicate experiment with 500-tip trees. Plots showing the point estimates for
all replicates can be found in
\cref{fig:gridptalpha,fig:gridptI,fig:gridptm,fig:gridptN}. For all parameters
except $m$, the error of point estimates was negatively correlated with the
number of sampled tips in the tree (for \gls{alpha}, \gls{I}, and \gls{N}
respectively: Spearman's $\rho$ = \Sexpr{tip.cor["alpha", round(estimate, 2)]},
\Sexpr{tip.cor["I", round(estimate, 2)]}, \Sexpr{tip.cor["N", round(estimate,
2)]}; $p$-values
    $\Sexpr{tip.cor["alpha", pp(p.value)]}$,
    $\Sexpr{tip.cor["I", pp(p.value)]}$,
    $\Sexpr{tip.cor["N", pp(p.value)]}$).
The 95\% highest density intervals obtained for all parameters were extremely
wide, occupying $>$75\% of the grid in all cases (\cref{fig:gridest}).

{\color{blue}\uline{Similar to the results obtained in the previous experiment,
the \gls{alpha} and \gls{I} parameters were estimated with comparably high
accuracy, while the other two parameters were less identifiable. Across all
replicates, $R^2$ values for the correlations between the estimated and true
values were
    $\Sexpr{tests["alpha", round(rsquared, 2)]}$,
    $\Sexpr{tests["I", round(rsquared, 2)]}$,
    $\Sexpr{tests["m", round(rsquared, 2)]}$, and
    $\Sexpr{tests["N", round(rsquared, 2)]}$
for \gls{alpha}, \gls{I}, \gls{m}, and \gls{N} respectively. The mean absolute
errors of the point estimates were
    $\Sexpr{tests["alpha", round(abs.err, 2)]}$,
    $\Sexpr{tests["I", round(abs.err)]}$,
    $\Sexpr{tests["m", round(abs.err, 2)]}$, and
    $\Sexpr{tests["N", round(abs.err)]}$,
representing
    $\Sexpr{tests["alpha", round(rel.err)]}$\%,
    $\Sexpr{tests["I", round(rel.err)]}$\%,
    $\Sexpr{tests["m", round(rel.err)]}$\%, and
    $\Sexpr{tests["N", round(rel.err)]}$\%
of the respective grids.}

\uline{Qualitatively, three of the parameters (excluding \gls{N}) exhibited
weak identifiability within particular sections of the grid. The \glspl{IQR}
for \gls{alpha} were nearly identical for the values \gls{alpha} $\leq 1$,
spanning approximately [0-1.3]. Similarly, the \glspl{IQR} for \gls{alpha} $>
1$ were all approximately [0.7-2]} (\cref{fig:gridest}). \uline{The
distributions of kernel scores for these groups of values were also
qualitatively similar} (\cref{fig:gridalpha}), \uline{although the value
\gls{alpha} = 1.25 exhibited a more pronounced peak in kernel scores around the
true value.}

\uline{For \gls{I}, the \glspl{IQR} were nearly identical for the values
\gls{I} $\leq 2000$ (approximately [500,3800]), and for the values \gls{I} $>
3000$ (approximately [1600-5000]). Kernel score distributions for all \gls{I}
values exibited a similar rounded shape} (\cref{fig:gridI}), \uline{although
this was most pronounced for values near the upper and lower extremes of the
grid.}

\uline{The \glspl{IQR} for \gls{m} were distinct for \gls{m} = 1 and 2, but
were similar for all values of \gls{m} $> 2$ (approximately [2-6]). The kernel
score distribution for \gls{m} = 1 was distinct from the others, having a large
peak at \gls{m} = 1, while the others exhibited a valley at \gls{m} = 1 and
were mostly flat everywhere else} (\cref{fig:gridm}).

\uline{The \glspl{IQR} for \gls{N} did not form any obvious groups like the
other parameters' \glspl{IQR}} (\cref{fig:gridest}). \uline{The kernel score
distribution for \gls{N} = 1000 was distinct, having a peak at the true value}
(\cref{fig:gridN}). \uline{The distributions for \gls{N} $> 1000$ were all
relatively similar, having a valley at the low end of the grid and being mostly 
flat everywhere else.}}

>>>>

{\color{red}\sout{The accuracy of grid search estimates largely paralleled that
of the \gls{kSVR} classifiers.}}\Cref{fig:gridest} shows point estimates and
{\color{blue}\uline{50\%}} and 95\% highest density intervals for each of the
\gls{BA} parameters, for one replicate experiment with 500-tip trees. Plots
showing the point estimates for all replicates can be found in
\cref{fig:gridptalpha,fig:gridptI,fig:gridptm,fig:gridptN}. For all parameters
except $m$, the error of point estimates was negatively correlated with the
number of sampled tips in the tree (for \gls{alpha}, \gls{I}, and \gls{N}
respectively: Spearman's $\rho$ = \Sexpr{tip.cor["alpha", round(estimate, 2)]},
\Sexpr{tip.cor["I", round(estimate, 2)]}, \Sexpr{tip.cor["N", round(estimate,
2)]}; $p$-values
    $\Sexpr{tip.cor["alpha", pp(p.value)]}$,
    $\Sexpr{tip.cor["I", pp(p.value)]}$,
    $\Sexpr{tip.cor["N", pp(p.value)]}$).
The 95\% highest density intervals obtained for all parameters were extremely
wide, occupying $>$75\% of the grid in all cases (\cref{fig:gridest}).

{\color{blue}\uline{Across all replicates, $R^2$ values for the correlations
between the estimated and true values were
    $\Sexpr{tests["alpha", round(rsquared, 2)]}$,
    $\Sexpr{tests["I", round(rsquared, 2)]}$,
    $\Sexpr{tests["m", round(rsquared, 2)]}$, and
    $\Sexpr{tests["N", round(rsquared, 2)]}$
for \gls{alpha}, \gls{I}, \gls{m}, and \gls{N} respectively. The mean absolute
errors of the point estimates were
    $\Sexpr{tests["alpha", round(abs.err, 2)]}$,
    $\Sexpr{tests["I", round(abs.err)]}$,
    $\Sexpr{tests["m", round(abs.err, 2)]}$, and
    $\Sexpr{tests["N", round(abs.err)]}$,
representing
    $\Sexpr{tests["alpha", round(rel.err)]}$\%,
    $\Sexpr{tests["I", round(rel.err)]}$\%,
    $\Sexpr{tests["m", round(rel.err)]}$\%, and
    $\Sexpr{tests["N", round(rel.err)]}$\%
of the respective grids.}

\uline{Qualitatively, each of the parameters exhibited weak identifiability
within particular sections of the grid} (\cref{fig:gridest}). \uline{The 50\%
\glspl{HDI} for \gls{alpha} were nearly identical for the values $\gls{alpha}
\leq 0.5$ (on average [X-X]) and for $\gls{alpha} \geq 1.5$ ([X-X]).
For \gls{I}, similar 50\% \gls{HPD} were observed for $\gls{I} \leq 1500$
(average [X-X]) and for $\gls{I} > 3000$ ([X-X]). For \gls{m}, both the 50\%
and 95\% \glspl{HDI} were generally identical for $\gls{m} > 2$ (average [X-X]
for 50\% and [X-X] for 95\% \gls{HDI}). We did not observe a similar pattern
for \gls{N}, however the boundaries of the \gls{HDI} did not always
monotonically increase with the actual value.}

\Cref{fig:gridalpha,fig:gridI,fig:gridm,fig:gridN} \uline{show kernel score
distributions of kernel scores along the grid for each parameter. The
distributions for some values, such as $\gls{alpha} = 1.25$, $\gls{I} = 500$
and $4500$, $\gls{m} = 1$, and $\gls{N} = 1000$, exhibited distinct peaks
around the true value. This indicates that these values produce distinctively
shaped trees that can be identified with the tree kernel, when the other
parameter values are fixed and known. However, for the majority of values of
each parameter, the score distributions were fairly flat around the true value. 
This means there is a range of values which produce similarly shaped trees, and
the parameter is less identifiable within that range.}}
